---
title: "01-carbon-emissions-and-food"
author: "Florence Galliers"
date: "17/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(extrafont)
#font_import()
library(countrycode)
library(plyr)
library(dplyr)

setwd("/Users/florentinagalliers/Google Drive/Harper/6 Data Visualisation & Analytics/Assignment/datavisualization/01-carbon-emissions")
```

## Introduction

This dataset comes from nu3 (www.nu3.de), a Germany based nutrition and health company.  It was complied for the #TidyTuesday data visualisation challenge by Kasia Kulma in February 2020. They obtained and analysed data from the Food and Agriculture Organisation. The dataset contains the annual CO2 emissions and consumption levels per person for 11 food categories in 130 countries (beef, pork, poultry, lamb/goat, fish, dairy products, wheat, rice, soybeans, nuts, eggs). It was created originally to compare the differences between plant based and animal-product based diets, but it gives a good overview of the consumption and emissions from these food categories throughout the world. 

```{r }
food_consumption <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-18/food_consumption.csv')

# How many countries?
unique(food_consumption$country)
# 130

# How many food categories?
unique(food_consumption$food_category)
# 11

# calculate total co2 emissions for each country
footprint_total <- food_consumption %>%
  dplyr::select(country, co2_emmission) %>%
  group_by(country) %>%
  dplyr::summarise(total = sum(co2_emmission))

# calculate total consumption for each country
consumption_total <- food_consumption %>%
  dplyr::select(country, consumption) %>%
  group_by(country) %>%
  dplyr::summarise(total = sum(consumption))

totals <- merge(footprint_total, consumption_total,
                by = "country")
colnames(totals) <- c("country", "footprint", "consumption")

totals$difference <- totals$footprint - totals$consumption
head(totals)

# Biggest differences between consumption and emissions, both positive and negative?
tail(sort(totals$difference), 5)
head(sort(totals$difference), 5)

top10 <- totals %>%
  filter(totals$difference >= (tail(sort(totals$difference), 5))[1] | 
           totals$difference <= (head(sort(totals$difference), 5))[5])

dotchart(sort(top10$difference), top10$country,
     pch = 16,
     col = "red")


```

Consumption per food category worldwide?
So which food category is most popular 

```{r}
# make category a factor
food_consumption$food_category <- factor(food_consumption$food_category)
levels(food_consumption$food_category)
# rename some of the levels
food_consumption$food_category <- revalue(food_consumption$food_category,
        c("Lamb & Goat" = "Lamb/Goat",
          "Milk - inc. cheese" = "Dairy",
          "Nuts inc. Peanut Butter" = "Nuts",
          "Wheat and Wheat Products" = "Wheat"))

# I want to colour the points by continent. so need to add a new column to the data called continent.

food_consumption$continent <- countrycode(food_consumption$country, 'country.name', 'continent')
continents <- unique(food_consumption$continent) # five continents here

# create colour palette of 5 colours

palette <- c("#ef476f", "#ffa600", "#06d6a0", "#118ab2", "#670067")

colour_frame <- data.frame(continents, palette)

food_plot <- merge(food_consumption, colour_frame,
      by.x = "continent",
      by.y = "continents",
      all.x = T)

food_plot$continent <- as.factor(food_plot$continent)

#names(food_plot)[names(food_plot) == "palette"] <- "pal"
# make colour palette
#levels(food_plot$continent)
#color_pallete_function <- colorRampPalette(
#  colors = c("#ef476f", "#ffd166", "#06d6a0", "#118ab2", "#073b4c"))
#num_colors <- nlevels(food_plot$continent)
#food_color_colors <- color_pallete_function(num_colors)

# remove 0 values to clean up plot
food_plot$consumption[which(food_plot$consumption == 0)] = NA
food_plot <- na.omit(food_plot)

food_plot %>%
  group_by(food_category, continent) 


# make object containing labels for boxplot, in the right order!
xlabels <- c("Dairy", "Wheat", "Rice", "Poultry", "Fish", "Pork", "Beef", "Eggs", "Nuts", "Lamb/Goat", "Soybeans")

# make food category a factor and set the order as above.
food_plot$food_category <- factor(food_plot$food_category, levels = c("Dairy", "Wheat", "Rice", "Poultry", "Fish", "Pork", "Beef", "Eggs", "Nuts", "Lamb/Goat", "Soybeans"), ordered = T)

boxplot(consumption ~ reorder(food_category, consumption),
        data = food_consumption)
# -----------------------------------------------------------------

png(filename = "consumption02.png",
    width = 28, height = 22, unit = "cm", res = 300)

set.seed(42) # so that the jitter function remains the same each time...

# Create starting boxplot of food consumption
boxplot(consumption ~ food_category, 
        # reorder categories in decreasing consumption level
        data = food_plot,
        outline = F, # remove outlying points
        xlim = c(0.5, 16.5), ylim = c(0, 350), # set x and y axis limits
        at = c(1, 2.5, 4, 5.5, 7, 8.5, 10, 11.5, 13, 14.5, 16),
        col = "grey95", medcol = "grey60", whiskcol = "grey60", # colours
        staplecol = "grey60", boxcol = "grey60", outcol = "grey60", #colours
        lwd = 0.5, # decrease line width to half
        ylab = "", yaxt = "none", # remove y axis and labels
        xlab = "", xaxt = "none", # remove x axis and labels
        frame = F,
        boxwex = 1.2) # remove frame box around plot

# add title to plot, and x and y axis labels
title("Dairy and wheat products are the most widely eaten food types worldwide",
      family = "Avenir",
      line = 0,
      adj = 0.5,
      cex = 2)

title(sub = "This boxplot shoes the consumption level of 11 food categories in 130 countries worldwide",
      family = "Avenir",
      cex = 0.5,
      col.sub = "grey40",
      line = -33,
      adj = 0.5)
      
title(ylab = "Consumption (kg/person/year)",
      xlab = "Food Category",
      family = "Avenir",
      line = 2.5)


# Insert y axis
axis(2, seq(0, 350, 50), # set labels and breaks
     las = 2, # rotate text to be horizontal
     cex.axis = 0.8, family = "Avenir",
     col = "grey60") # font size and style

# Insert x axis
axis(1, at = c(1, 2.5, 4, 5.5, 7, 8.5, 10, 11.5, 13, 14.5, 16), 
     labels = xlabels, # position and labels
     cex.axis = 0.9, family = "Avenir", # font size and style
     tick = F, # remove tick marks
     line = -1.6 # move x axis labels up towards plot
     )

# add strip charts for each continent
stripchart(consumption ~ food_category, 
           data = filter(food_plot, continent == "Africa"), 
           col = alpha(palette[1], 0.5), vertical = T,
           method = "jitter", add = T, 
           at = c(0.6, 2.1, 3.6, 5.1, 6.6, 8.1, 9.6, 11.1, 12.6, 14.1, 15.6), 
           pch = 16, ylim = c(0, 450))
stripchart(consumption ~ food_category, 
           data = filter(food_plot, continent == "Americas"), 
           col = alpha(palette[2], 0.5), # give points transparent effect
           vertical = T, method = "jitter", 
           add = T, # add to existing plot
           at = c(0.8, 2.3, 3.8, 5.3, 6.8, 8.3, 9.8, 11.3, 12.8, 14.3, 15.8), 
           # off-centre positions to get staggered effect
           pch = 16)
stripchart(consumption ~ food_category, 
           data = filter(food_plot, continent == "Asia"), 
           col = alpha(palette[3], 0.5), vertical = T,
           method = "jitter", add = T, 
           at = c(1, 2.5, 4, 5.5, 7, 8.5, 10, 11.5, 13, 14.5, 16), # positions in centre
           pch = 16)
stripchart(consumption ~ food_category, 
           data = filter(food_plot, continent == "Europe"), 
           col = alpha(palette[4], 0.5), vertical = T,
           method = "jitter", add = T, 
           at = c(1.2, 2.7, 4.2, 5.7, 7.2, 8.7, 10.2, 11.7, 13.2, 14.7, 16.2),
           pch = 16)
stripchart(consumption ~ food_category, 
           data = filter(food_plot, continent == "Oceania"), 
           col = alpha(palette[5], 0.5), vertical = T,
           method = "jitter", add = T, 
           at = c(1.4, 2.9, 4.4, 5.9, 7.4, 8.9, 10.4, 11.9, 13.4, 14.9, 16.4),
           pch = 16)

# add text annotations using expression and phantom strings for multicoloured
# Europe
text(x = 2.8, y = 300,
     expression("Europe "*phantom("has the highest")),
     col = palette[4],
     family = "Avenir",
     cex = 0.7)
text(x = 2.8, y = 300,
     expression(phantom("Europe ")*"has the highest"),
     col = "grey40",
     family = "Avenir",
     cex = 0.7)
text(x = 2.8, y = 289,
     "consumption levels of\ndairy products",
     col = "grey40",
     family = "Avenir",
     cex = 0.7)

# Asia 
text(x = 4.5, y = 217,
     "Rice consumption in",
     col = "grey40",
     family = "Avenir",
     cex = 0.7)
text(x = 4.5, y = 210,
     expression("many"*phantom(" Asian")*" countries"),
     col = "grey40",
     family = "Avenir",
     cex = 0.7)
text(x = 4.5, y = 210,
     expression(phantom("many")*" Asian"*phantom(" countries")),
     col = palette[3],
     family = "Avenir",
     cex = 0.7)
text(x = 4.5, y = 195,
     expression("is higher than on\nother continents"),
     col = "grey40",
     family = "Avenir",
     cex = 0.7)

# Africa
text(x = 7.5, y = 150,
     expression("Countries in "*phantom("Africa")*" consume the"),
     col = "grey40",
     family = "Avenir",
     cex = 0.7)
text(x = 7.5, y = 150,
     expression(phantom("Countries in ")*"Africa"*phantom(" consume the")),
     col = palette[1],
     family = "Avenir",
     cex = 0.7)
text(x = 7.5, y = 143,
     "lowest quantities of animal products",
     col = "grey40",
     family = "Avenir",
     cex = 0.7)

# Americas
text(x = 14, y = 80,
     "Animal products are consumed at",
     col = "grey40",
     family = "Avenir",
     cex = 0.7)
text(x = 14, y = 73,
     expression("higher levels in"*phantom(" America")),
     col = "grey40",
     family = "Avenir",
     cex = 0.7)
text(x = 14, y = 73,
     expression(phantom("higher levels in")*" America"),
     col = palette[2],
     family = "Avenir",
     cex = 0.7)

# Oceania

text(x = 10, y = 120,
     expression("Food consumption in"*phantom(" Oceania")*" is"),
     col = "grey40",
     family = "Avenir",
     cex = 0.7)
text(x = 10, y = 120,
     expression(phantom("Food consumption in")*" Oceania"*phantom(" is")),
     col = palette[5],
     family = "Avenir",
     cex = 0.7)
text(x = 10, y = 100,
     "consistently around the average
point. This lower variation may
be due to it being the smallest
continent",
     col = "grey40",
     family = "Avenir",
     cex = 0.7)

dev.off()


```

Battle of the animals
- Which meat is consumed most widely across the world, and which contributes the most CO2 emissions?

```{r}
# filter the dataset to only contain meat categories:
meat <- food_consumption %>%
  filter(food_category == "Pork" | food_category == "Poultry" | food_category == "Beef" | food_category == "Lamb/Goat" | food_category == "Fish")

# convert food categories to factors
meat$food_category <- as.factor(meat$food_category)

# test plot to see consumption vs category
ggplot(meat, 
       aes(x = reorder(food_category, consumption), 
           y = consumption)) +
  geom_bar(stat = "identity") +
  coord_flip()
  
# what about combining consumption and emissions onto one plot?
# first need to tidy data
meat_tidy <- meat %>%
  pivot_longer(names(meat)[3:4],
               names_to = "type",
               values_to = "kg")

# test plot to compare consumption and emissions of different types of meat
ggplot(meat_tidy, 
       aes(fill = type, x = food_category, y = kg)) +
  geom_bar(stat = "identity",
           position = "dodge") +
  coord_flip()
# can see there is a massive difference between the two
# looks fine but maybe using the 'difference' category would be better.
ggplot(meat, 
       aes(x = reorder(food_category, difference), 
           y = difference)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "",
       subtitle = "The difference between consumption of meat (kg/person/year) 
and their carbon emissions (kg CO2/person/year) worldwide")

# subset for each type of meat and find the highest value
beef <- meat %>%
  filter(food_category == "Beef")
beef[which.max(beef$co2_emmission), ]
argentina_value <- as.numeric(beef[which.max(beef$co2_emmission), 4])

lamb <- meat %>%
  filter(food_category == "Lamb/Goat")
lamb[which.max(lamb$co2_emmission), ]
iceland_value <- as.numeric(lamb[which.max(lamb$co2_emmission), 4])

fish <- meat %>%
  filter(food_category == "Fish")
fish[which.max(fish$co2_emmission), ]
maldives_value <- as.numeric(fish[which.max(fish$co2_emmission), 4])

pork <- meat %>%
  filter(food_category == "Pork")
pork[which.max(pork$co2_emmission), ]
hongkong_value <- as.numeric(pork[which.max(pork$co2_emmission), 4])

poultry <- meat %>%
  filter(food_category == "Poultry")
poultry[which.max(poultry$co2_emmission), ]
israel_value <- as.numeric(poultry[which.max(poultry$co2_emmission), 4])

# remove these values from the meat_edit dataset
meat_edit <- meat[-(which(meat$co2_emmission == argentina_value | 
        meat$co2_emmission == iceland_value |
        meat$co2_emmission == maldives_value |
        meat$co2_emmission == hongkong_value |
          meat$co2_emmission == israel_value)), ]

# subset each value into its own data frame so it can be plotted separately.
argentina <- meat %>%
  filter(country == "Argentina" & food_category == "Beef")
# Iceland - Lamb/Goat
iceland <- meat %>%
  filter(country == "Iceland" & food_category == "Lamb/Goat")
# Maldives - Fish
maldives <- meat %>%
  filter(country == "Maldives" & food_category == "Fish")
# Hong Kong - Pork
hongkong <- meat %>%
  filter(country == "Hong Kong SAR. China" & food_category == "Pork")
# Israel - Poultry
israel <- meat %>%
  filter(country == "Israel" & food_category == "Poultry")

# colour palette for 5 colours:
colpal <- c("#ff6361", "#58508d", "#ffa600", "#003f5c", "#bc5090")

# Plot a strip chart
consumption02 <- ggplot(meat_edit, 
       aes(x = reorder(food_category, co2_emmission), 
           y = co2_emmission, 
           size = consumption)) +
  geom_point(position = position_jitter(seed = 41),
             colour = alpha("grey50", 0.5)) +
  coord_flip() +
  labs(title = "Argentina emits the most CO2 from beef...",
       subtitle = "...which countries have the highest emissions from other animal products?",
       size = "Each circle represents a country,\nsized by quantity (kg/person/year) of\neach meat supplied for consumption") +
  ylab("CO2 emissions (kg CO2/person/year)") +
  scale_x_discrete(labels = c("POULTRY", "FISH", "PORK", "LAMB/GOAT", "BEEF")) +
  theme(plot.title = element_text(family = "Avenir", 
                                  size = 16, 
                                  colour = colpal[1]),
        plot.subtitle = element_text(family = "Avenir",
                                     size = 12),
        axis.title.y = element_blank(),
        axis.title.x = element_text(family = "Avenir",
                                    size = 8),
        plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill = "white"),
        panel.grid = element_blank(),
        axis.ticks = element_blank(),
        axis.text.y = element_text(angle = 0, hjust=1,
                                   size = 10, 
                                   family = "Avenir"),
        axis.text.x = element_text(family = "Avenir"),
        legend.position = c(0.85, 0.2),
        legend.text = element_text(family = "Avenir",
                                   size = 8),
        legend.title = element_text(family = "Avenir",
                                    size = 8),
        legend.background = element_blank(),
        legend.direction = "horizontal",
        plot.margin = unit(c(1,1,1,1), "cm")) +
  guides(size = guide_legend(reverse=TRUE,
                             title.position = "top", 
                             title.hjust = 0.5,
                             label.position = "bottom",
                             override.aes=list(colour = alpha("grey50", 0.5)))) +
  geom_point(data = argentina, position = position_jitter(seed = 41), 
             colour = colpal[1]) +
  geom_point(data = iceland, position = position_jitter(seed = 41), 
             colour = colpal[2]) +
  geom_point(data = hongkong, position = position_jitter(seed = 36), 
             colour = colpal[3]) +
  geom_point(data = maldives, position = position_jitter(seed = 41), 
             colour = colpal[4]) + 
  geom_point(data = israel, position = position_jitter(seed = 24), 
             colour = colpal[5]) +
  # argentina text
  annotate("text", x = 4.78, y = 1600,
           label = "Argentina",
           colour = colpal[1],
           size = 4,
           family = "Avenir") +
  # iceland text
  annotate("text", x = 3.77, y = 820,
           label = "Iceland",
           colour = colpal[2],
           size = 4,
           family = "Avenir") +
  # hong kong text
  annotate("text", x = 3.1, y = 360,
           label = "Hong Kong",
           colour = colpal[3],
           size = 4,
           family = "Avenir") +  
  # maldives text
  annotate("text", x = 1.77, y = 390,
           label = "Maldives",
           colour = colpal[4],
           size = 4,
           family = "Avenir") +
  # israel text
  annotate("text", x = 0.85, y = 140,
           label = "Israel",
           colour = colpal[5], 
           size = 4, 
           family = "Avenir")

consumption02

ggsave(filename = "consumption02.png", plot = consumption02,
       width = 24, height = 20, units = "cm")

```


```{r}
# subset dairy category into its own object
dairy <- food_consumption %>%
  filter(food_category == "Dairy") %>%
  dplyr::select(country, consumption)

library(sf)
library(viridis)
library(rnaturalearth)

world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)

world_dairy <- st_as_sf(merge(dairy, world,
                     by.x = "country",
                     by.y = "name",
                     all.x = TRUE))

europe_dairy <- world_dairy %>%
  subset(continent == "Europe")

europe_labels <- europe_dairy %>%
  dplyr::select(country, geometry, consumption) %>%
  filter(country != "Russia")

finland_label <- europe_labels %>%
  filter(country == "Finland")

belarus_label <- europe_labels %>%
  filter(country == "Belarus")


set.seed(41)

consumption03 <- ggplot(data = europe_dairy) +
  geom_sf(aes(fill = consumption), color = "grey30") +
  labs(title = "") +
  scale_fill_viridis(option = "D", na.value = "grey80",
                     direction = -1,
                     guide = guide_colourbar(),
                     breaks = c(150, 200, 250, 300, 350, 400),
                     labels = c("150kg/person/year", "200", "250", "300", "350", "400")) +
  coord_sf(xlim = c(-25.5, 32.87), ylim = c(30.15, 69.5), expand = FALSE,
           clip = "off") +
  theme_void() +
  ggrepel::geom_label_repel(data = europe_labels, 
                            aes(label = country, geometry = geometry),
                            stat = "sf_coordinates",
                size = 3,
                min.segment.length = 10,
                color = "grey25",
                family = "American Typewriter") +
  # add subtitle for finland
  geom_label(data = finland_label, 
               aes(label = country, fill = consumption), 
               x = -18, y = 55.5,
                size = 5,
             label.padding = unit(0.2, "cm"),
                color = "white",
             family = "American Typewriter") +
    geom_label(data = belarus_label, 
               aes(label = country, fill = consumption), 
               x = -18, y = 43.7,
                size = 5,
             label.padding = unit(0.2, "cm"),
                color = "gray30",
             family = "American Typewriter") +
  theme(
    legend.position = c(0.15, 0.5),
    legend.direction = "vertical",
    legend.title = element_blank(),
    legend.text = element_text(family = "American Typewriter",
                               size = 7),
    plot.margin = unit(c(1,1,1,1), "cm")
  ) +
  annotate("text",
           x = -18,
           y = 60,
           label = "Who consumes the most\ndairy products?",
           family = "American Typewriter",
           size = 7) +
  guides(fill = guide_colorbar(barwidth = 1, 
                               barheight = 10))

consumption03

ggsave(filename = "consumption03.png", plot = consumption03,
       width = 28, height = 24, unit = "cm")

#mypalette <- colorNumeric( palette="magma",
#                           domain=world_dairy$consumption, na.color="transparent")

#leaflet(world_dairy) %>%
#  addTiles() %>%
#  addPolygons(fillColor = ~mypalette(consumption), stroke = F)


```

