---
title: "03-malaria"
author: "Florence Galliers"
date: "17/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# set up and install packages
library(tidyverse)
library(dplyr)
library(ggplot2)
library(viridis)
library(rnaturalearth)
library(countrycode)
library(GGally)
library(doBy)
library(ggrepel)
library(tidyr)
library(ggtext)

setwd("/Users/florentinagalliers/Google Drive/Harper/6 Data Visualisation & Analytics/Assignment/datavisualization/03-malaria")

```

This set of visualisations have been created using three different but related datasets about malaria incidence and mortality rates around the world. The datasets were created by ‘Our World in Data’ (https://ourworldindata.org/malaria). They also formed part of the #TidyTuesday data visualisation challenge in December 2018. The datasets contain the following data:
-	Malaria incidence by country for all ages in 2000, 2005, 2010 and 2015.
-	Annual malaria death rate by country for all ages between 1990 and 2016.
-	Annual total malaria deaths by country, split into 5 age groups, between 1990 and 2016.


```{r}
# Get data
tuesdata <- tidytuesdayR::tt_load('2018-11-13')
malaria_deaths_age <- tuesdata$malaria_deaths_age
malaria_deaths <- tuesdata$malaria_deaths
malaria_inc <- tuesdata$malaria_inc

# Explore the datasets
head(malaria_deaths_age)
# 6 columns
# X1, entity = country, code, year, age group and number of deaths
# There are 30780 observations

# How many countries?
unique(malaria_deaths_age$code) # 197
unique(malaria_deaths_age$entity) #228
unique(malaria_deaths_age$age_group)
# this should be a factor? and ordered
malaria_deaths_age$age_group <- factor(malaria_deaths_age$age_group, 
                                       levels = c("Under 5",
                                                  "5-14",
                                                  "15-49",
                                                  "50-69",
                                                  "70 or older"))

unique(malaria_deaths_age$year)
# year goes from 1990 to 2016
```

Malaria 02 - ggplot

```{r}
# Lets go and look at incidence. 2000, 2005, 2010, 2015
# Parallel plot of African countries over time
head(malaria_inc)
# contains incidence data as a rate, per 1,000 of population at risk.

# Extract just countries in continent of Africa from this?
# create new column that contains the continent using country code package
malaria_inc$continent <- countrycode(malaria_inc$Entity, 
            "country.name",
            "continent")
# then subset to just africa
africa_inc <- malaria_inc %>%
  subset(continent == "Africa")

# check it has worked
head(africa_inc)
# rename ridiculously long column name for incidence
africa_inc <- africa_inc %>%
  rename(incidence = 'Incidence of malaria (per 1,000 population at risk) (per 1,000 population at risk)')

# subset three countries that have the highest incidence rate in 2015
africa_inc_high <- africa_inc %>%
  subset(Year == 2015) %>%
  arrange(desc(incidence)) %>%
  slice(1:3)

africa_inc_high <- africa_inc %>%
  filter(Entity == "Mali" | Entity == "Burkina Faso" | Entity == "Nigeria")

# subset the three countries with the largest difference between 2015 and 2000 incidence rates, e.g. those that have decreased their malaria incidence the most
africa_inc_2000 <- africa_inc %>%
  subset(Year == 2000)

africa_inc_2015 <- africa_inc %>%
  subset(Year == 2015)

africa_inc_diff <- africa_inc_2000 %>%
  select(Entity) %>%
  mutate(diff = (africa_inc_2000$incidence - africa_inc_2015$incidence)) %>%
  arrange(desc(diff)) %>%
  slice(1:3)

# cant workout how to filter directly using the above... just use character strings here but not ideal incase dataset was ever updated...
africa_inc_low <- africa_inc %>%
  filter(Entity == "Ethiopia" | Entity == "Guinea-Bissau" | Entity == "Sao Tome and Principe")

# make a plot of incidence over time
ggplot(africa_inc, aes(x = Year, 
                       y = incidence,
                       group = Entity)) + # group by country
  geom_line(color = "grey90") + # make all lines light grey
  geom_line(data = africa_inc_high, color = "#C12910", size = 1) + # red for high incidence
  geom_line(data = africa_inc_low, color = "#2A9D8F", size = 1) + # green for low incidence
  coord_trans(clip = "off") + # so labels dont get cut off
  theme_minimal() + 
  # add text for high incidence country labels
  geom_text(data = subset(africa_inc_high, Year == 2010), 
            aes(label = Entity, y = incidence + 15),
            color = "#C12910",
            hjust = 0.5, x = 2010,
            family = "Futura Medium") +
  # add text for low incidence country labels
  geom_text(data = subset(africa_inc_low, Year == 2015), 
            aes(label = Entity, y = incidence-10),
            color = "#2A9D8F", x = 2015,
            family = "Futura Medium") +
  labs(title = "There is a still vast disparity in Malaria incidence rates throughout Africa",
       y = "Incidence Rate (per 1,000 people)") +
  theme(
    panel.grid = element_blank(), # remove gridlines
    plot.title = element_text( family = "Futura Medium", size = 16),
    axis.title = element_text(family = "Futura Medium", size = 9, hjust = 0.5),
    axis.text = element_text(family = "Futura Medium", size = 10)) 

# Add another element? 
# subset death data for africa
# create new column for continent as before on the incidence data above
malaria_deaths$continent <- countrycode(malaria_deaths$Entity, 
            "country.name",
            "continent")
# subset for africa, and for 2015
africa_deaths <- malaria_deaths %>%
  subset(continent == "Africa") %>%
  filter(Year == 2015)
# check data
head(africa_deaths)
# rename long column name to deaths
africa_deaths <- africa_deaths %>%
  rename(deaths = `Deaths - Malaria - Sex: Both - Age: Age-standardized (Rate) (per 100,000 people)`)

# merge incidence data with death data
africa_total <- merge(africa_inc, africa_deaths,
                      by.x = "Entity",
                      by.y = "Entity",
                      all.x = T) %>%
  dplyr::select(Entity, Year.x, incidence, deaths) %>%
  rename("year" = "Year.x")

# make totals of the highest incidence rates
africa_total_high <- africa_total %>%
  filter(Entity == "Mali" | Entity == "Burkina Faso" | Entity == "Nigeria")

# and largest decrease in incidence rate
africa_total_low <- africa_total %>%
  filter(Entity == "Ethiopia" | Entity == "Guinea-Bissau" | Entity == "Sao Tome and Principe")

# rename Sao Tome and Principe so it will appear on two lines in the label
africa_total_low$Entity <- gsub("Sao Tome and Principe", "Sao Tome\nand Principe", africa_total_low$Entity)

africa_total_low

# plot and make line widths show the number of deaths
malaria02 <- ggplot(africa_total, aes(x = year, 
                       y = incidence,
                       group = Entity, 
                       size = deaths)) + # size lines by 2015 deaths
  geom_line(color = "grey90", alpha = 0.6) + # background lines slightly transparent
  # add lines for high incidence countries, varying the colors 
  geom_line(data = africa_total_high, color = c("#C12910","#C12910","#C12910","#C12910",
                                                "#980000", "#980000", "#980000", "#980000",
                                                "#FB5E3C", "#FB5E3C", "#FB5E3C", "#FB5E3C"),
            aes(size = deaths), alpha = 0.6, # slight transparency
            show.legend = F) + # remove legend
  # add lines for largest decrease in incidence countries
  geom_line(data = africa_total_low, color = c("#2A9D8F", "#2A9D8F", "#2A9D8F", "#2A9D8F",
                                               "#007362", "#007362", "#007362", "#007362",
                                               "#64CFBF", "#64CFBF", "#64CFBF", "#64CFBF"),
            aes(size = deaths), alpha = 0.6,
            show.legend = F) +
  coord_trans(clip = "off") + # remove clipping so labels don't get cut off
  theme_minimal() +
  geom_label_repel(data = subset(africa_total_high, year == 2015), 
            aes(label = Entity, y = incidence),
            seed = 42, # set seed as geom_label_repel is random each time
            color = c("#C12910", "#980000", "#FB5E3C"), # same colours as lines
            hjust = "left", # left justification
            x = 2015.1, # slightly move over from end of lines
            family = "Futura Medium", size = 4,
            nudge_x = 0.2,
            min.segment.length = Inf, # remove any spider lines
            xlim = c(2015.2, NA)) +
  geom_label_repel(data = subset(africa_total_low, year == 2015), 
            aes(label = Entity, y = incidence),
            seed = 42,
            color = c("#2A9D8F", "#007362", "#64CFBF"), 
            hjust = "left",
            x = 2015.1,
            family = "Futura Medium", size = 4, 
            min.segment.length = Inf,
            xlim = c(2015.2, NA)) +
  labs(title = "There is vast disparity in the rate of Malaria incidence throughout Africa",
       subtitle = "Despite some countries <span style = 'color:#007362;'>drastically reducing</span> the prevalence of malaria over the last 15 years, other <br>countries still have a yearly malaria incidence rate of <span style = 'color:#980000;'>over 400 people per 1,000</span> at risk.",
       y = "Incidence Rate (per 1,000 people)",
       x = "Year",
       # change title of legend
       size = "Each line represents one country in\nAfrica, with its width representing the\ndeath rate per 100,000 people in 2015") +
  theme(
    panel.grid = element_blank(), # remove gridlines
    # titles and subtitles
    plot.title = element_markdown(family = "Futura Medium", size = 16),
    plot.subtitle = element_markdown(family = "Futura Medium", size = 12, color = "grey40"),
    axis.title = element_text(family = "Futura Medium", size = 10, hjust = 0.5, color = "grey40"),
    axis.text = element_text(family = "Futura Medium", size = 10, color = "grey40"),
    # legend
    legend.position = c(0.92, 0.87),
    legend.title = element_text(family = "Futura Medium", size = 10, color = "grey40"),
    legend.text = element_text(family = "Futura Medium", size = 10, color = "grey40"),
    legend.key.width = unit(2, "cm"),
    # margins
    plot.margin = unit(c(1, 3, 1, 1), "cm"),
    axis.text.y = element_text(hjust = 2),
    axis.text.x = element_text(vjust = 3)
    ) 

malaria02

ggsave(filename = "malaria02.png", plot = malaria02, 
       width = 28, height = 20, unit = "cm")

```

Malaria 01 - Base R Graphics

```{r}
# look at data
head(malaria_deaths_age)

# make age group into a factor
malaria_deaths_age$age_group <- factor(malaria_deaths_age$age_group, 
                                       levels = c("Under 5",
                                                  "5-14",
                                                  "15-49",
                                                  "50-69",
                                                  "70 or older"))

# summarise all worldwide deaths by year and by age group

# first check entries are only those that are countries, and not internal summaries:
unique(malaria_deaths_age$entity)
# there is already an Entity called 'World'
world_deaths_age <- malaria_deaths_age %>%
  filter(entity == "World")

# make summaries of age group for each year
world_death_summary <- world_deaths_age %>%
  group_by(age_group, year) %>%
  summarise(total = sum(deaths))

# change into wide format in order to make barplot in base R... not tidy data ahh!
world_death_summary_wide <- world_death_summary %>%
  pivot_wider(names_from = year,
              values_from = total)

# transform into matrix
world_death_summary_wide <- as.matrix(world_death_summary_wide)
# remove first column as it is still the factors
world_death_summary_wide <- world_death_summary_wide[ ,-1]
# change row names to factor names
rownames(world_death_summary_wide) <- c("Under 5", "5-14", "15-49", "50-69", "70+")

# make a barplot
barplot(world_death_summary_wide,
        col = c("darkred", "blue", "green", "orange", "yellow"), 
        space = 0.2)

# -----------------------------------------

# make barplot for Low SDI
low_sdi_deaths <- malaria_deaths_age %>%
  filter(entity == "Low SDI")

# make summaries of age group for each year
low_sdi_summary <- low_sdi_deaths %>%
  group_by(age_group, year) %>%
  summarise(total = sum(deaths))

# change into wide format in order to make barplot in base R... not tidy data ahh!
low_sdi_summary_wide <- low_sdi_summary %>%
  pivot_wider(names_from = year,
              values_from = total)

# transform into matrix
low_sdi_summary_wide <- as.matrix(low_sdi_summary_wide)
# remove first column as it is still the factors
low_sdi_summary_wide <- low_sdi_summary_wide[ ,-1]
# change row names to factor names
rownames(low_sdi_summary_wide) <- c("Under 5", "5-14", "15-49", "50-69", "70+")

# make a barplot
barplot(low_sdi_summary_wide,
        col = c("darkred", "blue", "green", "orange", "yellow"), 
        space = 0.2)

# ---------------------------------------


# make barplot for high SDI
high_sdi_deaths <- malaria_deaths_age %>%
  filter(entity == "High SDI")

# make summaries of age group for each year
high_sdi_summary <- high_sdi_deaths %>%
  group_by(age_group, year) %>%
  summarise(total = sum(deaths))

# change into wide format in order to make barplot in base R... not tidy data ahh!
high_sdi_summary_wide <- high_sdi_summary %>%
  pivot_wider(names_from = year,
              values_from = total)

# transform into matrix
high_sdi_summary_wide <- as.matrix(high_sdi_summary_wide)
# remove first column as it is still the factors
high_sdi_summary_wide <- high_sdi_summary_wide[ ,-1]
# change row names to factor names
rownames(high_sdi_summary_wide) <- c("Under 5", "5-14", "15-49", "50-69", "70+")

# make a barplot
barplot(high_sdi_summary_wide,
        col = c("darkred", "blue", "green", "orange", "yellow"), 
        space = 0.2)

# -----------------------------------------------


# make barplot for middle SDI
middle_sdi_deaths <- malaria_deaths_age %>%
  filter(entity == "Middle SDI")

# make summaries of age group for each year
middle_sdi_summary <- middle_sdi_deaths %>%
  group_by(age_group, year) %>%
  summarise(total = sum(deaths))

# change into wide format in order to make barplot in base R... not tidy data ahh!
middle_sdi_summary_wide <- middle_sdi_summary %>%
  pivot_wider(names_from = year,
              values_from = total)

# transform into matrix
middle_sdi_summary_wide <- as.matrix(middle_sdi_summary_wide)
# remove first column as it is still the factors
middle_sdi_summary_wide <- middle_sdi_summary_wide[ ,-1]
# change row names to factor names
rownames(middle_sdi_summary_wide) <- c("Under 5", "5-14", "15-49", "50-69", "70+")

# make a barplot
barplot(middle_sdi_summary_wide,
        col = c("darkred", "blue", "green", "orange", "yellow"), 
        space = 0.2)


# -----------------------------------

# make a comparison barplot between low, middle and high SDI countries.
palette <- c("#fc8d62", "#66c2a5", "#8da0cb", "#e78ac3", "#a6d854")

age_groups <- c("Under 5", "5-14", "15-49", "50-69", "70+")

years <- seq(1990, 2016, 1)

# write plot into png file, 28cm x 20cm, resolution 300 dpi.
png(filename = "malaria01.png",
    width = 28, height = 20, units = "cm", res = 300)

# place on right hand side
par(fig = c(0.5, 0.99, 0.5, 1))
# make a barplot for low sdi countries
barplot(low_sdi_summary_wide,
        col = palette, 
        ylim = c(0, 550000), 
        space = 0.3,
        yaxt = "none",
        las = 2, cex.names = 0.6, line = -0.8,
        family = "Avenir",
        border = palette)
# add ylab title
title(ylab = "Number of Deaths from Malaria per Year",
      family = "Avenir",
      cex.lab = 0.8, 
      line = 3)
# add xlab title
title(xlab = "Year",
      family = "Avenir",
      cex.lab = 0.8,
      line = 1.5)
# y axis
axis(2, las = 2, 
     at = c(0, 50000, 100000, 150000, 
            200000, 250000, 300000, 
            350000, 400000, 450000, 
            500000, 550000),
     labels = c("0", "50,000", "100,000",
                "150,000", "200,000", "250,000",
                "300,000", "350,000", "400,000",
                "450,000", "500,000", "550,000"),
     cex.axis = 0.5,
     cex = 0.5,
     family = "Avenir",
     tck = -0.01, gap.axis = 0.5,
     lwd = 0.5)
# plot title
title("Low SDI",
      family = "Avenir", 
      cex.main = 1.2)

# change position to the bottom left
par(fig = c(0.01, 0.5, 0, 0.5), new = T)

# make a barplot for middle sdi countries
barplot(middle_sdi_summary_wide,
        col = palette, 
        space = 0.3,
        ylim = c(0, 20000),
        yaxt = "none",
        las = 2, cex.names = 0.6,
        line = -0.8,
        family = "Avenir",
        border = palette)
title(ylab = "Number of Deaths from Malaria per Year",
      family = "Avenir",
      cex.lab = 0.8, 
      line = 3)
title(xlab = "Year",
      family = "Avenir",
      cex.lab = 0.8,
      line = 1.5)
axis(2, las = 2, at = c(0, 5000, 10000, 15000, 20000),
     labels = c("0", "5000", "10000", "15000", "20000"),
     cex.axis = 0.5, lwd = 0.5, 
     cex = 0.5, tck = -0.01,
     family = "Avenir")
title("Middle SDI",
      family = "Avenir", 
      cex.main = 1.2,
      col.main = "grey20")

# change position to the bottom right
par(fig = c(0.5, 0.99, 0, 0.5), new = T)
# make a barplot for high sdi countries
barplot(high_sdi_summary_wide,
        col = palette, 
        space = 0.3,
        ylim = c(0, 15),
        yaxt = "none",
        las = 2, cex.names = 0.6,
        line = -0.8,
        family = "Avenir",
        border = palette)
title(ylab = "Number of Deaths from Malaria per Year",
      family = "Avenir",
      cex.lab = 0.8, 
      line = 3)
title(xlab = "Year",
      family = "Avenir",
      cex.lab = 0.8,
      line = 1.5)
axis(2, las = 2, at = c(0, 5, 10, 15),
     labels = c("0", "5", "10", "15"),
     cex.axis = 0.5, lwd = 0.5,
     cex = 0.5, tck = -0.01,
     family = "Avenir")
title("High SDI",
      family = "Avenir", 
      cex.main = 1.2,
      col.main = "grey20")

# add titles and text to top left of plot
par(fig = c(0, 0.5, 0.5, 1))

# add legend
#legend("bottom", title = "Age Group",
#       legend = c("Under 5", "5-14", "15-49", "50-69", "70 or older"),
#       fill = palette, border = palette,
#       cex = 0.3, horiz = T, pt.cex = 1,
#       bty = "n")

# add rectangles and text for legend instead because the legend functionality in Base R is not doing what I want it to when messing with the par() settings
rect(4, 3, 8, 4, 
     col = palette[1])
rect(9, 3, 13, 4, 
     col = palette[2])
rect(14, 3, 18, 4, 
     col = palette[3])
rect(19, 3, 23, 4, 
     col = palette[4])
rect(24, 3, 28, 4, 
     col = palette[5])

text(6, 3.5, label = age_groups[1],
     family = "Avenir", cex = 0.6)
text(11, 3.5, label = age_groups[2],
     family = "Avenir", cex = 0.6)
text(16, 3.5, label = age_groups[3],
     family = "Avenir", cex = 0.6)
text(21, 3.5, label = age_groups[4],
     family = "Avenir", cex = 0.6)
text(26, 3.5, label = age_groups[5],
     family = "Avenir", cex = 0.6)
# add label on left of the 'legend' rectangles
text(1.5, 3.5, label = "Age Group:", 
     family = "Avenir", cex = 0.6)
# add title
text(x = 16, y = 13, 
     "Childhood deaths from malaria are 300,000x\nhigher in countries that are less developed",
     family = "Avenir",
     cex = 1.12)
# add subtitle
text(x = 16, y = 8, 
     "Depending on a countries socio-demographic index (SDI)\nthere is a huge difference between the number of deaths,\nthe distribution of these deaths between age groups and the\nrate of decrease over time. The SDI is a scale that identifies\na countries position on the spectrum of development.",
     family = "Avenir",
     cex = 0.7)
# add note
text(x = 16, y = 1,
     "Please note the extreme differences in scales of the\nnumber of deaths per year for the three graphs",
     family = "Avenir",
     cex = 0.7)

# end plot
dev.off()


```

Malaria 04 - Circle Packing Plot

```{r}

devtools::install_github("jeromefroe/circlepackeR")
library(circlepackeR)
library(data.tree)
library(d3treeR)
library(packcircles)
library(ggplot2)
library(viridis)
library(ggiraph)
library(tidyverse)

# subset incidence data to just 2015
malaria_2015 <- malaria_inc %>%
  filter(Year == 2015)

# make new column for continent
malaria_2015$continent <- countrycode(malaria_2015$Entity, "country.name", "continent")
# remove NA values
malaria_2015 <- malaria_2015 %>%
  na.omit()
# reduce column name (this is just making a new column...)
malaria_2015$incidence <-  malaria_2015$`Incidence of malaria (per 1,000 population at risk) (per 1,000 population at risk)`
# check data is looking okay
head(malaria_2015)
# subset only the needed columns
malaria <- malaria_2015[ ,-c(2,3,4)]
head(malaria)

# remove any 0 values
malaria <- malaria %>%
  filter(incidence > 0)

data <- malaria

# Add a column with the text you want to display for each bubble:
data$text <- paste("Country: ", data$Entity, "\n",
                   "Incidence Rate:", data$incidence, "\n",
                   "Continent: ", data$continent, "\n" )

# Generate the layout
packing <- circleProgressiveLayout(data$incidence, sizetype='area')
data <- cbind(data, packing)
dat.gg <- circleLayoutVertices(packing, npoints = 50)

# 50 points for each circle so 51 repetitions here
dat.gg$value <- rep(data$incidence, each = 51)

# style settings for tooltip popups
tooltip_css <- "color:white;padding:10px;font-family: Avenir;border-radius:5px;"

# make the plot
plot_malaria <- ggplot() + 
  geom_polygon_interactive(data = dat.gg, 
                           aes(x, y, 
                               group = id, 
                               fill = -value, # make highest incidence the darkest colouts
                               tooltip = data$text[id], 
                               data_id = id), 
                           colour = "black", 
                           alpha = 0.6) +
  scale_fill_viridis() +
  geom_text(data = data, 
            aes(x, y, 
                label = gsub("Country: ", "", Entity),
                size = incidence), 
            color="black",
            family = "Avenir") +
  scale_size_continuous(range = c(1,3)) +
  theme_void() + 
  labs(title = "Malaria Incidence Rates in 2015") +
  theme(legend.position = "none", 
        plot.margin = unit(c(0,0,0,0),"cm"),
        plot.title = element_text(
          size = 20,
          family = "Avenir"
        )) + 
  coord_equal() 

# Turn it interactive
int_malaria <- girafe(ggobj = plot_malaria, 
                fonts = "Avenir",
                options = list(
                  opts_tooltip(use_fill = T, css = tooltip_css),
                  opts_hover_inv(css = "opacity:0.4;")))
print(int_malaria)

```

