---
title: "02-meteorites"
author: "Florence Galliers"
date: "17/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(extrafont)
library(showtext)
#font_import()

library(maps)
library(countrycode)
library(devtools)
devtools::install_github('rensa/ggflags')
library(ggflags)
library(tidyr)

setwd("/Users/florentinagalliers/Google Drive/Harper/6 Data Visualisation & Analytics/Assignment/datavisualization/02-meteorites")
```

Meteorite Data Set contains information about where and when meteorites were found or fell. There is a difference between a 'found' meteorite and a 'fell' meteorite. Those that are classified as 'fell' are those which were seen falling from the sky and later tracked down successfully, whereas those which were 'found' do not have a exact date or location of their falling.

```{r}
meteorites <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-11/meteorites.csv")

# start by summarising the count of meteorites per year
meteorites_total <- meteorites %>%
  group_by(year) %>%
  dplyr::summarise(Count = n())
# plot
plot(x = meteorites_total$year,
     y = meteorites_total$Count,
     type = "l")

# subset data for 1900 --> present
which.max(meteorites_total$year)
meteorites_total[265,] # this is the year 2101? in the future... 
meteorites_total <- meteorites_total[-265, ]
which.max(meteorites_total$year)
meteorites_total[264,] # the most recent is now 2013.

# subset data 1900-2013
meteorites_total <- meteorites_total %>%
  subset(year > 1900)
# plot again
plot(x = meteorites_total$year,
     y = meteorites_total$Count,
     type = "l")

# still not that much happening up until 1970 - lets subset again
meteorites_total <- meteorites_total %>%
  subset(year > 1970)
# plot again
plot(x = meteorites_total$year,
     y = meteorites_total$Count,
     type = "l")

# what about just fallen meteorites because those that were found may not have fallen in that year.
meteorites_total2 <- meteorites %>%
  filter(fall == "Fell") %>%
  group_by(year) %>%
  dplyr::summarise(Count = n())

plot(x = meteorites_total2$year,
     y = meteorites_total2$Count,
     type = "l")

# again not much going on until 1750? so lets subset down
meteorites_total2 <- meteorites_total2 %>%
  subset(year > 1750)

plot(x = meteorites_total2$year,
     y = meteorites_total2$Count,
     type = "l",
     col = "black",
     lwd = 2)

# group by decade
meteorites3 <- meteorites %>% 
  mutate(decade = year - year%%10) %>% 
  filter(decade>=1750, decade<=2005) %>% 
  select(name,mass,fall, year,decade, lat,long) %>% 
  drop_na()

# filter again to just fallen meteorites
meteorites_total3 <- meteorites3 %>%
  filter(fall == "Fell") %>%
  group_by(decade) %>%
  dplyr::summarise(Count = n())

```

```{r}
# Downloaded title font from (https://www.dafont.com/galaxy-1.font)

# make objects containing text
subtitle <- "Fallen meteorites are those that were spotted falling and later located"

section1 <- "The idea of\nmeteorites was\nlargely rejected\nbefore the 1790s"

section2 <- "Scientific knowlegde of meteorites increased\nduring the 19th century, as well as the world\npopulation and education levels."

section3 <- "The number of meteorites has\nrecently stablised, there may\nalso be a time lag between a\nmeteorite falling and its\nappearance in records."

png(filename = "meteorites01.png", width = 30, height = 20, units = "cm", res = 300)

# plot
par(bg = "white", xpd = NA)

plot(x = meteorites_total3$decade,
     y = meteorites_total3$Count,
     type = "l",
     col = "grey30",
     lwd = 4,
     xlab = "",
     ylab = "",
     bty = "n",
     xaxt = "none", yaxt = "none",
     cex.lab = 0.8, 
     col.lab = "grey50")

# add title
title("The Rise of Fallen Meteorites",
      family = "Galaxy 1", 
      cex.main = 3,
      col.main = "grey20")

# add y axis title
title(ylab = "No. of observed falls (av. over 10 year period)",
      line = 1.2, 
      col.lab = "grey50",
      cex.lab = 0.8,
      family = "Courier")

title(xlab = "Year",
      line = 1.5, 
      col.lab = "grey50",
      cex.lab = 0.8,
      family = "Courier")

# add subtitle
text(1915, 92, subtitle,
     col = "grey30",
     family = "Courier")

# x axis
axis(1, seq(1750, 2000, 10),
     labels = F,
     col = "grey50",
     col.axis = "grey50",
     tck = -0.005)

text(x = seq(1750, 2000, 10),
     y = par("usr")[3] - 2,
     labels = meteorites_total3$decade,
     xpd = NA, srt = 35,
     adj = 0.9, cex = 0.8,
     col = "grey50",
     family = "Courier")

# y axis
axis(2, seq(0, 90, 10),
     las = 2,
     col = "grey50",
     col.axis = "grey50",
     cex.axis = 0.8,
     tck = -0.01,
     line = -0.8,
     family = "Courier")

# add rectangle for first section
rect(1750, 0, 1785, 90,
     col = alpha("#99D6EA", 0.5),
     border = "white")
# add text to first section
text(1767, 15, section1,
     col = "grey30",
     cex = 0.8,
     family = "Courier")

# add rect for second section
rect(1786, 0, 1930, 90,
     col = alpha("#FDD85D", 0.5),
     border = "white")
# add text to second section
text(1850, 80, section2,
     col = "grey30",
     cex = 0.9,
     family = "Courier")

# add rect for third section
rect(1931, 0, 2000, 90,
     col = alpha("#6798C0", 0.5),
     border = "white")
# add text to third section
text(1970, 80, section3,
     col = "grey30",
     cex = 0.8,
     family = "Courier")

# add lines for important events
segments(1794, 0, 1794, 12,
      col = "coral",
      lwd = 2)
segments(1833, 0, 1833, 23,
      col = "coral",
      lwd = 2)

# add text for important events
text(1812, 5, "1794: Ernst Chladni,\n'father of meteoritics'\nhypothesised that\nmeteorites are rocks\nfrom space",
     col = "coral",
     family = "Courier",
     cex = 0.6)

text(1855, 10,
     "1833: Leonid Shower,\nsparked modern day\nawareness of meteorites\nand after this point\nscientific studies began",
     col = "coral",
     family = "Courier",
     cex = 0.6)

lines(x = meteorites_total3$decade,
     y = meteorites_total3$Count,
     type = "l",
     col = "grey30",
     lwd = 4,
     col.lab = "grey50")

# add points
points(x = meteorites_total3$decade,
       y = meteorites_total3$Count,
       pch = 8, 
       cex = 1.5,
       col = "#6798C0")

dev.off()

```

Now for a plot using ggplot

```{r}

# order the meteorites by their mass
meteorites_ordered <- meteorites[order(-meteorites$mass), ]

# find the top say 20 largest? and subset
top20 <- meteorites_ordered[1:20,]
# Convert mass from g to tonnes
top20$mass <- (top20$mass/1000000)

# plot mass as a bar chart, reordering names by mass
ggplot(top20,
       aes(x = reorder(name, mass),
           y = mass,
           fill = class)) +
  geom_bar(stat = "identity") +
  coord_flip() + # flip axis so names down the side
  labs(y = "Mass (tonnes)",
       x = "")

# get the countries for each of these top 20 meteorites.
country <- map.where(database = "world",
                     top20$long, # use lat and long values from dataset
                     top20$lat)

country # entry 2 and 16 are NAs

top20$geolocation[2] # Greenland
country[2] <- "Greenland"
top20$geolocation[16] # Brazil
country[16] <- "Brazil"

# use country code to convert country name into iso2 code ready to use with ggflag
country_code <- countrycode(country, 
            "country.name",
            "iso2c")

# change to lower case
country_code <- tolower(country_code)

# bind country code with top 20 
top20c <- cbind(top20, country_code)

# combine some classes
top20c$class <- factor(top20c$class)
levels(top20c$class)
levels(top20c$class) <- c("Stony Chondrites",
                          "Iron (Metallic)",
                          "Iron (Metallic)",
                          "Iron (Metallic)",
                          "Iron (Metallic)",
                          "Iron (Metallic)",
                          "Iron (Metallic)",
                          "Iron (Metallic)",
                          "Iron (Metallic)",
                          "Iron (Metallic)",
                          "Iron (Metallic)",
                          "Stony-Iron",
                          "Stony-Iron")
top20c$class <- factor(top20c$class, levels = c("Iron (Metallic)", "Stony-Iron", "Stony Chondrites")) 

# create new column combining name of meteorite and year it fell
top20c$title <- paste(top20c[ , 1],",", top20c[ , 7])

# now replot including flags
meteorites02 <- ggplot(top20c,
       aes(x = reorder(title, mass), y = mass, fill = class)) +
  geom_segment(aes(x = reorder(title, mass), xend = title, 
                   y = 0, yend = mass,
                   color = class),
               size = 3, alpha = 0.2) +
  geom_segment(aes(x = reorder(title, mass), xend = title, 
                   y = 0, yend = mass,
                   color = class),
               size = 1, alpha = 0.8) +
  geom_flag(aes(x = reorder(title, mass), y = 0, 
                country = country_code), 
            size = 6) + 
  geom_label(label = round(top20$mass, 0),
             nudge_y = 1,
             colour = "white",
             family = "Avenir",
             show.legend = F) +
  geom_label(label = "Mass in tonnes",
             x = 8.5, y = 48,
             colour = "white",
             fill = "grey60",
             family = "Avenir") +
  scale_color_manual(values = c("#8B8682", "coral", "#ffcc00")) +
  scale_fill_manual(values = c("#8B8682",  "coral", "#ffcc00")) +
  coord_flip(clip = "off") +
  labs(title = "Name of Meteorite,   Year                     Country                                                                                                                                                                                                                                    Mass in Tonnes",
       y = "",
       x = "") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(hjust = 0,
                               colour = "#615d5b",
                               family = "Avenir", 
                               size = 10),
    plot.background = element_rect(fill = "white"),
    plot.title = element_text(colour = "#8B8682",
                              family = "Avenir",
                              size = 8,
                              vjust = 0,
                              hjust = 1),
    panel.grid = element_blank(),
    plot.margin = unit(c(1, 1, 1, 1), "cm"),
    legend.title = element_blank(),
    legend.position = c(0.77, 0.33),
    axis.title.x = element_text(colour = "#615d5b",
                                family = "Avenir",
                                size = 10),
    axis.text.x = element_blank(),
    axis.line.x = element_blank()
  ) +
  annotate("text",
           x = 12, 
           y = 48, 
           label = "The largest meteorites found\non earth are made of Iron",
           color = "#615d5b",
           family = "Avenir",
           size = 7) +
  annotate("text",
           x = 10, 
           y = 48,
           label = "18 out of 20 of the largest meteorites that\nhave been found or fallen were made of Iron.",
           colour = "#8B8682",
           family = "Avenir",
           size = 4) 

# save file as meteorites02
ggsave(filename = "meteorites02.png", plot = meteorites02, 
       width = 28, height = 20, unit = "cm")


```

Interactive plot: where are the meteorites in the world?

```{r}

library(leaflet)
library(forcats)
library(leaflet.extras)

meteorites$class <- factor(meteorites$class)
unique(meteorites$class)

stony_chondrites_carbonaceous <- c("CR2-an", "CI1", "CV3", "CM2", "OC", "C2-ung",
                                   "CO3.3", "CBa", "CR2", "CO3.2", "CK4", "CO3.5", 
                                   "CO3.6", "C3-ung", "CO3.4", "C", "CO3.7", "CH3",
                                   "CO3", "CM1/2", "CM1", "CO3.0", "CV3-an", "CK6",
                                   "CK5", "C6", "C4", "C1-ung", "CK3", "C4-ung",
                                   "CK4/5", "CK3-an", "CR", "CM-an", "CK5/6", "C2",
                                   "C4/5", "CR1", "Relict OC", "CBb", "CH/CBb", 
                                   "CO3.8", "CK4-an", "CB", "C5/6-ung", "CV2", 
                                   "CK", "OC3", "CK3/4", "CO3.1", "CK3.8", "CM2-an", 
                                   "C3.0-ung", "CR7", "CM", "C3/4-ung", "CR-an", "C1/2-ung")

stony_chondrites_ordinary <- c("L5", "H6", "L6", "LL3-6", "H5", "L", "H4", "H",
                               "LL5", "L/LL4", "L3", "LL6", "L4", "LL4", "LL3.8", 
                               "LL3.15", "LL3.9", "H/L3.9", "H4-an", "L/LL6", "H/L4",
                               "H4-5", "L3.7", "LL3.4", "H3.8", "H3-5", "H5-6", "H5-7",
                               "L3-6", "H4-6", "H3", "L/LL5", "H3.7", "H4/5", "H3/4", "H?", 
                               "H3-6", "L3.4", "L3.7-6", "L5/6", "L3.6", "LL3.2", "L3-4", 
                               "H5/6", "L5-6", "LL3.6", "H3-4", "LL", "L4-6", "LL3.00" ,
                               "H3.4", "LL3.3", "H/L3.6", "LL7", "LL4-6", "L/LL6-an", 
                               "H3.9/4", "L3.8", "LL5-6", "LL3.8-6", "L3.9", "L4-5", 
                               "L3-5", "LL4/5", "L4/5", "H3.9", "H3.6-6", "H3.8-5",
                               "H3.8/4", "H3.9-5", "L3.9/4", "H~5", "H~6", "L/LL3.10",
                               "LL3", "L~6", "L~3", "H~4", "L(LL)3.5-3.7", "H3.6", "L3.4-3.7",
                               "L3.5", "LL7(?)", "LL6(?)", "H3.5-4", "H3.5", "L/LL3",
                               "H7", "L(?)3", "L3.2", "L3.7-3.9", "LL3.7", "LL3.5",
                               "L3.7-4", "L3.3", "L3.1", "H3.3", "H3.7-6", "H3.0", "H3.1",
                               "L3.0", "L/LL3.4", "LL3.0", "H5-an", "L6/7", "Relict H", 
                               "L-imp melt", "H3-an", "L~5", "H3.9-6",  "LL~6", "L3.9-6",
                               "H-an", "L/LL3-6", "L/LL3-5", "H/L3.5", "H/L3", "LL4-5",
                               "H/L6", "L3/4", "H-imp melt", "H(L)3-an", "L(LL)3",
                               "H(L)3", "L7", "L/LL~6", "L/LL~5", "L~4", "L/LL~4",
                               "LL(L)3 ", "H3.2", "L-melt breccia", "H6-melt breccia",
                               "H5-melt breccia", "H-melt rock", "LL5/6", "LL3/4", 
                               "H3.4/3.5", "H/L5", "H(5?)", "LL-imp melt", "H~4/5",
                               "L6-melt breccia", "L3.5-3.7", "L3.3-3.7", "L3.2-3.6",
                               "L3.3-3.6", "L3.05",  "L/LL3.2", "L3.5-5", "L/LL(?)3", 
                               "H4(?)", "L5-7", "L-melt rock", "H-metal", "L-metal",
                               "L/LL5/6", "H3.8-4", "H/L~4",  "H4/6", "H3.7-5", "LL3.7-6",
                               "H3.7/3.8", "L3.7/3.8", "LL-melt rock", "L3.5-3.9", 
                               "L3.2-3.5", "L3.3-3.5", "L3.0-3.7", "L(LL)3.05", "L3.10",
                                "L/LL4/5", "L3.8-an", "L3.8-6", "LL3.10", "L3.8-5", 
                               "H/L3-4", "L(H)3", "LL6/7", "LL3.1", "LL~4", "LL~4/5", 
                               "L(LL)~4", "H3.05", "H3.10", "LL3-5", "H/L3.7", 
                               "LL3-4", "LL<3.5", "L/LL-melt rock", "H6/7", "L4-melt rock",
                               "L6-melt rock", "H/L4/5", "H/L6-melt rock", "L3-7", "LL3.05",
                               "LL4/6", "LL3.8-4", "H3.15", "LL-melt breccia", "LL6-melt breccia",
                               "L5-melt breccia", "LL(L)3.1", "LL6-an", "L4-melt breccia", 
                               "H4-melt breccia", "L3-melt breccia", "L~4-6", "LL~5", 
                               "H-melt breccia", "L3.00", "H3.0-3.4", "L/LL4-6", 
                               "L4-an", "H3.8/3.9", "L3.9-5", "H3.8-6", "H3.4-5", 
                               "L3.0-3.9", "L3.5-3.8", "H3.2-3.7", "L3.6-4", "L/LL3.5",
                               "L/LL3.6/3.7", "H/L4-5", "LL~3", "LL5-7", "LL3.9/4", "H3.8-an",
                               "L/LL5-6", "L(LL)5", "L(LL)6", "LL3.1-3.5", "H3.2-6", 
                               "H(?)4", "H3.2-an", "L/LL", "LL(L)3")

stony_chondrites_enstatite <- c("EH4", "EL6", "EH3/4-an", "EH7-an", "EH3",  "EH5",
                                "E4", "EL5", "EL3", "E6", "EH4/5", "EL4", "EL4/5",
                                "E", "EH", "EL6/7", "EL7", "EH-imp melt", "E3-an", "E3",
                                "EH6", "EL-melt rock", "EL3/4", "EH7", "E-an", "E5",
                                "E5-an", "EH6-an")

stony_chondrites <- c("Stone-uncl", "K3", "R3.8-6", "R3.8-5", "Chondrite-ung", "R3.6",
                      "R4", "R3-6", "R3.8", "R3.5-6", "R3-4", "Chondrite-fusion crust", 
                      "R3", "R5", "R", "Fusion crust", "R6", "K", "R3/4", "R3.9", "R4/5",
                      "R3.7", "R3-5", "R3.4", "R3.5-4")

stony_primitive_achondrites <- c("Acapulcoite", "Ureilite-an", "Ureilite",  "Lodranite",
                                 "Winonaite", "Brachinite",  "Ureilite-pmict",
                                 "Acapulcoite/Lodranite", "Acapulcoite/lodranite", 
                                 "Achondrite-prim", "Lodranite-an")

stony_achondrites <- c("Diogenite-pm", "Eucrite-mmict", "Angrite", "Aubrite", "Howardite",
                       "Eucrite-pmict", "Diogenite", "Eucrite", "Martian (chassignite)",
                       "Eucrite-cm", "Martian (nakhlite)", "Eucrite-br", "Martian (shergottite)",
                       "Achondrite-ung", "Martian (OPX)", "Eucrite-Mg rich", "Eucrite-unbr", 
                       "Lunar (anorth)", "Lunar", "Lunar (gabbro)", "Lunar (feldsp. breccia)",
                       "Eucrite-an", "Lunar (bas/anor)", "Lunar (basalt)", "Diogenite-an", 
                       "Aubrite-an", "Diogenite-olivine", "Lunar (bas. breccia)", "Enst achon",
                       "Impact melt breccia", "Martian", "Lunar (bas/gab brec)", "Enst achon-ung",
                       "Howardite-an", "Martian (basaltic breccia)", "Lunar (norite)", 
                       "Stone-ung")

stony_iron_achondrites <- c("Mesosiderite-A1", "Mesosiderite", "Mesosiderite-A3/4",
                            "Pallasite", "Mesosiderite-A3", "Pallasite, PMG", "Mesosiderite-B2",
                            "Pallasite, PMG-an", "Mesosiderite-B1", "Mesosiderite-B4",
                            "Mesosiderite-an", "Mesosiderite-A2", "Pallasite, PES",
                            "Mesosiderite-A", "Mesosiderite?",  "Mesosiderite-B",
                            "Mesosiderite-A4", "Mesosiderite-C2", "Pallasite, ungrouped",
                            "Mesosiderite-C", "Pallasite?")

iron_achondrites <- c("Iron, IVA", "Iron, IIAB", "Iron, IAB-sLL", "Iron, ungrouped",
                      "Iron, IAB-MG", "Iron?", "Iron, IIIAB", "Iron, IID", "Iron, IIE",
                      "Iron, IAB-sHL", "Iron", "Iron, IIE-an", "Iron, IAB-ung", "Iron, IIF",
                      "Iron, IIIAB-an", "Iron, IAB complex", "Iron, IIIE-an",  "Iron, IAB-sLM",
                      "Iron, IC", "Iron, IID-an", "Iron, IIIE", "Iron, IVA-an", "Iron, IIG",
                      "Iron, IIC", "Iron, IIIF", "Iron, IVB", "Iron, IAB?", "Iron, IAB-sLH",
                      "Iron, IIAB-an", "Iron, IIIAB?", "Iron, IAB-sHH", "Relict iron", 
                      "Iron, IAB-an", "Iron, IC-an", "Iron, IAB-sHL-an", "Iron, IIE?")

other_unknown <- c("Unknown")


meteorites_class <- fct_collapse(meteorites$class,
             scc = stony_chondrites_carbonaceous,
             sco = stony_chondrites_ordinary,
             sce = stony_chondrites_enstatite,
             sc = stony_chondrites,
             spa = stony_primitive_achondrites,
             sa = stony_achondrites,
             sia = stony_iron_achondrites,
             ia = iron_achondrites,
             ou = other_unknown)

meteorites$class <- as.factor(meteorites_class)

levels(meteorites_class)

labels <- c("Stony Primitive Achondrites",
            "Stony Achondrites",
            "Stony Chondrites - Carbonaceous",
            "Stony Chondrites",
            "Stony Chondrites - Enstatite",
            "Stony Chondrites - Ordinary",
            "Iron Achondrites",
            "Stony Iron Achondrites",
            "Other / Unknown")

# create bins for color palette
bins <- c(0, 20, 40, 60, 80, 120, 150, Inf)
# create palette
palette <- colorFactor("Paired", domain = meteorites$class)
    
mytext <- paste(meteorites$name,
                    "<br/>",
                    meteorites$year,
                    "<br/>",
                    "Size (kg): ", (meteorites$mass)/1000,
                    "<br/>",
                    sep="") %>%
      lapply(htmltools::HTML)
    
leaflet(meteorites) %>% 
  addProviderTiles(providers$CartoDB.DarkMatter) %>%
  setView(lat = 20, lng = 0, zoom = 1.6) %>%
  addCircleMarkers(~long, ~lat, # Points for circles
             fillColor = ~palette(class), 
             fillOpacity = 0.7, # make opaque
             color = "white",  # colour of edge?
             radius = ~log(mass), # radius
             stroke = FALSE, # no edge on circles
             weight = 1,# hmm edge of lines?
             clusterOptions = markerClusterOptions(
               disableClusteringAtZoom = 4,
               spiderfyOnMaxZoom = F
             ),
    label = mytext, 
    labelOptions = labelOptions(style = list("font-weight" = "normal",
                                             padding = "3px 8px"), 
                                textsize = "13px", 
                                direction = "auto")) %>%
  addLegend(position = "bottomright",
            pal = palette,
            values = ~class,
            title = "Class of Meteorite",
            opacity = 1,
            labFormat = function(type, cuts, p) {  # Here's the trick
                                                  paste0(labels)
                             }
                             ) 

```

Sankey diagram

```{r}

library(networkD3)
library(devtools)
#install_github("Displayr/flipPlots") #install from github
library(flipPlots)
library(plyr)
library(dplyr)

# remove zero values
zero_values <- which(meteorites$mass == 0)
meteorites_all <- meteorites[-zero_values, ]
# remove NAs for mass
meteorites_all <- meteorites %>%
  na.omit()

# how many meteorites have fallen?
fallen <- meteorites_all %>%
  filter(fall == "Fell") %>%
  dplyr::summarise(Count = n())

# how many meteorites have been found?
found <- meteorites_all %>%
  filter(fall == "Found") %>%
  dplyr::summarise(Count = n())

summary(meteorites_all$mass)
hist(meteorites_all$mass, breaks = 10, freq = T)

# how many meteorites in each class
meteorites_all_class <- fct_collapse(meteorites_all$class,
             scc = stony_chondrites_carbonaceous,
             sco = stony_chondrites_ordinary,
             sce = stony_chondrites_enstatite,
             sc = stony_chondrites,
             spa = stony_primitive_achondrites,
             sa = stony_achondrites,
             sia = stony_iron_achondrites,
             ia = iron_achondrites,
             ou = other_unknown)

meteorites_all$class <- as.factor(meteorites_all_class)

levels(meteorites_all$class)

meteorites_all$class <- revalue(meteorites_all$class,
        c("scc" = "Stony Chondrites - Carbonaceous",
          "sco" = "Stony Chondrites - Ordinary",
          "sce" = "Stony Chondrites - Enstatite",
          "sc" = "Stony Chondrites",
          "spa" = "Stony Primitive Achondrites",
          "sa" = "Stony Achondrites",
          "sia" = "Stony-Iron Achondrites",
          "ia" = "Iron Achondrites",
          "ou" = "Other / Unknown"))

meteorites_all$fall <- as.factor(meteorites_all$fall)

meteorites_all$masscat <- cut(meteorites_all$mass, c(-1, 49, 99, 999, 9999, Inf),
                              labels = c("0-49g", "50-99g", "100g-999g", "1kg-9.999kg", "10kg-60 tonnes"))

classes <- meteorites_all %>%
  group_by(fall, masscat, class) %>%
  dplyr::summarise(Count = n())

colourpalette <- c("grey80", "grey60",
                   "#AED6F1", "#5DADE2", "#3498DB", "#2E86C1", "#2874A6",
                   "red", "purple", "orange", "pink", "green", "darkgreen", "yellow", "coral")

SankeyDiagram(classes[ ,-4],
              link.color = "Source",
              weights = classes$Count, 
              label.show.counts = TRUE,
              label.show.varname = F,
              font.family = "Avenir",
              hovertext.show.percentages = T,
              sinks.right = F,
              colors = colourpalette)

```

